\documentclass{article}
\usepackage[utf8]{inputenc}

\title{\textit{Lymphocryptovirus human gammaherpesvirus 4:} Epstein-Barr Virus}
\author{Kai Drumm}
\date{January 2023}

\begin{document}

\maketitle

\section{At A Glance}

\subsection{Morphology and size}

DNA inside of a round protein nucleocapsid, surrounded loosely by tegument and membrane envelope with glycoproteins. This virus is 122-180 nm diameter.

\subsection{Taxonomic data}

Herpesviridae is a the family containing all herpes viruses. There are three subfamilies: 
\begin{itemize}
    \item Alphaherpesvirinae ($\alpha$) 
    \item Betaherpesvirinae ($\beta$)
    \item Gammaherpesvirinae ($\gamma$)
\end{itemize}

\noindent EBV is in the Gammaherpesvirinae subfamily. \newline

There are nine herpes viruses that infect humans, accross all three subfamilies:
\begin{enumerate}
    \item HSV-1, oral herpes ($\alpha$)
    \item HSV-2, genital herpes ($\alpha$)
    \item Varicella zoster ($\alpha$)
    \item Epstein-Barr ($\gamma$)
    \item Cytomegalovirus ($\beta$)
    \item Roseolovirus (two types) ($\beta$)
    \item Kaposki's sarcoma-associated herpesvirus ($\gamma$)
\end{enumerate}

\subsection{Reservoir}

EBV is common in humans, with 90\% of adults showing evidence of past or current infection. It uses a latency strategy to persist in the body despite immune resistance, expressing a limited subset of its genes. Bone marrow may be one site of persistence.

\subsection{Mode of transmission}

Saliva, sexual contact

\subsection{Natural history of disease}

Infection is often asymptomatic in childhood. In adolescence and adulthood, it can cause infectious mononucleosis and other diseases.

\subsection{Means of diagnosis}

EBV can be diagnosed by detecting antibodies to it or the small RNAs that it produces.

\section{History and Burden}

\subsection{History of discovery}

EBV was discovered through research on Burkitt's Lymphoma in Uganda in 1961-1968. Michael Epstein was a researcher who took the samples from Uganda to be studied in London and published the results along with his PhD student, Yvonne Barr. Denis Burkitt was a surgeon practicing in Uganda who presented information about the lymphoma.

\subsection{History of burden}

\subsection{Current global burden and distribution}

EBV is one of the most common viruses in humans. It increases the incidence of cancers, chronic fatigue, and multiple sclerosis.

\section{Countermeasures}

\section{Modeling}

\end{document}
